---
title: "WGCATCH 2025, ToR c BYC inventory"
author: "Karolina, Sarah"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r}
# PACKAGES
library(tidyverse, quietly = TRUE)
library(DT)
library(readxl)
```

```{r}
# READ data 

# New datacall 
path <- "../ToRc/Data/"
d2 <- read.csv(paste0(path, "D2.csv"))
d4 <- read.csv(paste0(path, "D4.csv"))

# Old datacall 
d2old <- read_xlsx(paste0(path, "eu.2022.04b_dataset.xlsx"), sheet = "Coverage")
## Remove first row - 
d2old <- d2old[-1, ]
## Rename colnames 
names(d2old) <- c(
  "Country",
  "Year",
  "SamplingScheme",
  "Gear",
  "Area",
  "NoTripsSampled",
  "NoTripsTotal",
  "SamplingIntensity",
  "ImportanceBirds",
  "ImportanceMammals",
  "ImportancePetFish",
  "ImportanceElasmobranchs",
  "ImportanceReptiles",
  "Comments"
)

d4old <- read_xlsx(paste0(path, "eu.2022.04b_dataset.xlsx"), sheet = "Coverage")

# RDBES - CE 
ce <- read.csv(paste0(path, "CommercialEffort.csv"))
## Filter out countries that did not respond to the datacall 
# unique(d2$Country)
# unique(ce$CEvesselFlagCountry)

ce <- filter(ce, CEvesselFlagCountry %in% unique(d2$Country))
# setdiff(unique(d2$Country), unique(ce$CEvesselFlagCountry)) # med countries and PT-20 excl - correct
```


```{r}
# Useful fun 
# helper to match a code as a standalone-ish token (prevents partial hits)
tok <- function(x) paste0("(?<![A-Z])(", x, ")(?![A-Z])")
# Clasify gears
# Group gears 
classify_gear <- function(x) {
  cats <- c(
    PelTrawl = str_detect(x, regex(tok("OTM|PTM|PT"), ignore_case = TRUE)),
    DemTrawl = str_detect(x, regex(tok("OTB|PTB|TBB|OTT|TB"), ignore_case = TRUE)),
    DemSeine    = str_detect(x, regex(tok("SSC|SPR|SDN"), ignore_case = TRUE)),
    PelSeine    = str_detect(x, regex(tok("PS"), ignore_case = TRUE)),
    LinesHooks = str_detect(x, regex(tok("LLD|LLS|LH|LHP|LHM|LTL"), ignore_case = TRUE)),
    Gillnetters = str_detect(x, regex(tok("GNS|GTR|GN|GND|GTN|GNC"), ignore_case = TRUE))
  )

  # the free text thing
  cats["PelTrawl"]   <- cats["PelTrawl"]   | str_detect(x, regex("pelagic trawler", ignore_case = TRUE))
  cats["Gillnetters"]<- cats["Gillnetters"]| str_detect(x, regex("gillnetter", ignore_case = TRUE))
  cats["DemTrawl"]   <- cats["DemTrawl"]   | (str_detect(x, regex("trawler", ignore_case = TRUE)) &
                                               str_detect(x, regex("demersal|cod", ignore_case = TRUE)))

  present <- names(cats)[cats]

  if (length(present) == 0) {
    "Others"
  } else if (length(present) == 1) {
    present
  } else {
    paste0("Multiple (", paste(sort(present), collapse = "+"), ")")
  }
}

# Expand area
expandIcesArea <- function(x) {
  x <- str_replace_all(x, " ", "")
  x <- str_replace(x, "\\.$", "")   # drop dot

  # Pure 27 region or fully specified already
  if (x == "27") return(x)
  if (str_detect(x, "^27\\.[0-9]+(\\.[A-Za-z0-9]+)*$")) return(x)

  # 27.4b, 27.7bcjk 
  if (str_detect(x, "^27\\.[0-9]+[A-Za-z]+$")) {
    num     <- str_match(x, "^27\\.([0-9]+)")[, 2]
    letters <- str_match(x, "^27\\.[0-9]+([A-Za-z]+)$")[, 2]
    letters_vec <- strsplit(letters, "")[[1]]
    letters_vec <- letters_vec[letters_vec != ""]
    return(sprintf("27.%s.%s", num, letters_vec))
  }

  # 7bcjk, 7fgh etc 
  if (str_detect(x, "^[0-9]+[A-Za-z]+$")) {
    num     <- str_match(x, "^([0-9]+)")[, 2]
    letters <- str_match(x, "^[0-9]+([A-Za-z]+)$")[, 2]
    letters_vec <- strsplit(letters, "")[[1]]
    letters_vec <- letters_vec[letters_vec != ""]
    return(sprintf("27.%s.%s", num, letters_vec))
  }

  # 27.1, 27.12 etc keep as is will be expanded
  if (str_detect(x, "^27\\.\\d+$")) return(x)

  # Fallback: keep unchanged
  x
}

# Parse
parse_one_code <- function(code) {
  orig <- code

  code <- str_squish(code)

  if (is.na(code) || code == "" || code == "NA") {
    return(tibble(
      original       = orig,
      AreaCode_system = NA_character_,
      AreaCode        = NA_character_
    ))
  }

  # Split on both "~" and "," (mixed strings)
  parts <- str_split(code, "[~,]")[[1]] %>%
    str_squish()
  
  parts <- parts[parts != ""]

  map_dfr(parts, function(p) {
    if (is.na(p) || p == "") {
      return(tibble(
        original        = orig,
        AreaCode_system = NA_character_,
        AreaCode        = NA_character_
      ))
    }

    # --- GSA: explicit GSA + number ----------------------------------------
    if (str_detect(str_to_upper(p), "^GSA\\s*\\d+$")) {
      num <- as.integer(str_extract(p, "\\d+"))
      return(tibble(
        original        = orig,
        AreaCode_system = "GSA",
        AreaCode        = paste0("GSA ", num)
      ))
    }

    # Plain digits
    if (str_detect(p, "^\\d{1,2}$")) {
      # If original string has a 27.* and no 'GSA'  interpret as 27.<n> (like 27.12,14)
      if (str_detect(orig, "27\\.") &&
          !str_detect(orig, regex("\\bGSA\\b", ignore_case = TRUE))) {
        new_code <- paste0("27.", p)
        return(tibble(
          original        = orig,
          AreaCode_system = "ICES",
          AreaCode        = new_code
        ))
      } else {
        # Otherwise treat as GSA number
        num <- as.integer(p)
        return(tibble(
          original        = orig,
          AreaCode_system = "GSA",
          AreaCode        = paste0("GSA ", num)
        ))
      }
    }

    p_clean <- str_replace_all(p, " ", "")
    p_clean <- str_replace(p_clean, "\\.$", "")

    if (str_detect(p_clean, "^27(\\.|$)") ||
        str_detect(p_clean, "^[0-9]+[A-Za-z]+$")) {
      expanded <- expandIcesArea(p_clean)
      # expanded may be length > 1
      return(tibble(
        original        = orig,
        AreaCode_system = "ICES",
        AreaCode        = expanded
      ))
    }

    # Unknown stuff 
    tibble(
      original        = orig,
      AreaCode_system = "Unknown",
      AreaCode        = p
    )
  })
}

```

```{r}
# Bind the d2 
d2old <- mutate(d2old, NoTripsSampled = as.numeric(NoTripsSampled),
                NoTripsTotal = as.numeric(NoTripsTotal))
d2all<- bind_rows(d2, d2old)

# Bind the d4
```



# Data sources 

Three different data sources were used in this analysis. First, the countries responses following the 2025 datacall, the 2020 datacall and the CE table from the RDBES dataset. 


# Data cleaning 

## Gear/Metier groupings 

The metiers reported by each country were grouped into five categories: 

1. Pelagic trawl
2. Demersal trawl
3. Demersal seine
4. Pelagic seine
5. Lines and hooks
6. Others 

The table below shows how metiers were grouped to the above categories. 



```{r}


# Group metiers - new datacall

d2allg <- mutate(d2all, GearGroup = vapply(Gear, classify_gear, character(1)))
tabgear <- distinct(d2allg, Gear, GearGroup)
datatable(tabgear, caption = "Grouped metiers in the new datacall")

# Group metiers - old datacall 
# 
# d2old <- mutate(d2old, GearGroup = vapply(metier_lvl, classify_gear, character(1)))
# # TODO - This needs to be checked!!! Function does not capture everything 
# tabgearold <- distinct(d2old, metier_lvl, GearGroup)
# datatable(tabgearold, caption = "Grouped metiers in the old datacall")

# Group gears - CE - run at the end its taking too long
# ce <- mutate(ce, GearGroup = vapply(CEmetier6, classify_gear, character(1))) # takes too long
# tabgearce <- distinct(ce, CEmetier6, GearGroup)
# datatable(tabgearce, caption = "Grouped metiers in the old datacall")
```


```{r}
# Look up table 
ices_divisions <- tribble(
  ~subarea,   ~division,
  # 27.1
  "27.1",     "27.1.a",
  "27.1",     "27.1.b",
  # 27.2
  "27.2",     "27.2.a",
  "27.2",     "27.2.b",
  # 27.3 (Baltic + Kattegat)
  "27.3.a",   "27.3.a.20",
  "27.3.a",   "27.3.a.21",
  "27.3.b",   "27.3.b.23",
  "27.3.c",   "27.3.c.22",
  # 27.3.d (Baltic Proper) incl. 28.1/28.2
  "27.3.d",   "27.3.d.24",
  "27.3.d",   "27.3.d.25",
  "27.3.d",   "27.3.d.26",
  "27.3.d",   "27.3.d.27",
  "27.3.d",   "27.3.d.28",
  "27.3.d",   "27.3.d.28.1",
  "27.3.d",   "27.3.d.28.2",
  "27.3.d",   "27.3.d.29",
  "27.3.d",   "27.3.d.30",
  "27.3.d",   "27.3.d.31",
  "27.3.d",   "27.3.d.32",
  # 27.4 (North Sea)
  "27.4",     "27.4.a",
  "27.4",     "27.4.b",
  "27.4",     "27.4.c",
  # 27.5 (Iceland & Faroes)
  "27.5",     "27.5.a",
  "27.5",     "27.5.b",
  # 27.6 (W of Scotland/Rockall)
  "27.6",     "27.6.a",
  "27.6",     "27.6.b",
  "27.6.b",   "27.6.b.1",
  # 27.7 (Celtic/Biscay)
  "27.7",     "27.7.a",
  "27.7",     "27.7.b",
  "27.7",     "27.7.c",
  "27.7",     "27.7.d",
  "27.7",     "27.7.e",
  "27.7",     "27.7.f",
  "27.7",     "27.7.g",
  "27.7",     "27.7.h",
  "27.7",     "27.7.j",
  # 27.8 (Bay of Biscay)
  "27.8",     "27.8.a",
  "27.8",     "27.8.b",
  "27.8",     "27.8.c",
  "27.8",     "27.8.d",
  # 27.9
  "27.9",     "27.9.a",
  # 27.10 (Azores)
  "27.10",    "27.10.a",
  "27.10.a",  "27.10.a.1",
  "27.10.a",  "27.10.a.2",
  # 27.12 (Svalbard/Barents fringe)
  "27.12",    "27.12.a",
  "27.12",    "27.12.b",
  # 27.14 (Greenland)
  "27.14",    "27.14.a",
  "27.14",    "27.14.b"
)
```





```{r}

# Collapse list cols, keep original row

d2e <- d2allg %>%
  ungroup() %>%
  mutate(.orig_row = row_number()) %>%
  rowwise() %>%
  mutate(
    AreaCode_collapsed = paste(AreaCode, collapse = "~")
  ) %>%
  ungroup()

# Expand general
d2_expanded <- d2e %>%
select(-AreaCode) %>%
  mutate(parsed = map(AreaCode_collapsed, parse_one_code)) %>% # the error is in the nested functions above, not here
  unnest(parsed) %>%
  # clean up
  select(-AreaCode_collapsed)

# Expand subarea -> subdiv 
d2div <- d2_expanded %>%
  mutate(
    is_ices_subarea  = str_detect(AreaCode, "^27\\.\\d+$"),
    is_ices_division = str_detect(AreaCode, "^27\\.\\d+\\.[a-z]")   # already division
  ) %>%
  left_join(ices_divisions, by = c("AreaCode" = "subarea")) %>%
  mutate(
    AreaCode1 = case_when(
      is_ices_subarea ~ coalesce(division, AreaCode),
      TRUE            ~ AreaCode
    )
  ) %>%
  select(-division, -is_ices_subarea, -is_ices_division) %>%
  distinct(.orig_row, AreaCode, AreaCode_system, .keep_all = TRUE)


```



```{r}
# Fix specific stuff 
d2div <- mutate(d2div, AreaCode = ifelse(AreaCode %in% "37.2.1", "GSA 17", AreaCode))
# Exclude 3271 France (talked with Sebastien)
# Exclude 27.3.bc 
d2div <- filter(d2div, !(AreaCode %in% "GSA 3721" & Country %in% "FR"))
d2div <- filter(d2div, !(AreaCode %in% "27.3.bc" & Country %in% "FR"))
# Reassign 3.a to 4.b (rectangle ovelap issue)
d2div <- mutate(d2div, AreaCode = ifelse(Country %in% "FR" & AreaCode %in% "27.3.a", "27.4.b", AreaCode))

# Exclude Finland area 27.3.d. (all Baltic) - Email for new submission
d2div <- filter(d2div, !AreaCode %in% "27.3.d.")

# Fix Slovenia 
d2div <- mutate(d2div, AreaCode_system = ifelse(Country %in% "SI", "GSA", AreaCode_system))
```





```{r}
d2div <- d2div %>%
  mutate(AreaCode = ifelse(substr(AreaCode, 1, 4) == "27.3", substr(AreaCode, 1, 9), sub("^(([^.]*\\.){2}[^.]*).*", "\\1", AreaCode)))
# d2divold<- d2divold %>%
#   mutate(AreaCode = ifelse(substr(AreaCode, 1, 4) == "27.3", substr(AreaCode, 1, 9), sub("^(([^.]*\\.){2}[^.]*).*", "\\1", AreaCode)))
# 
# # change naems, keep necessary fields and R bind
# 
# d2O <- select(d2divold, country, year, scheme_name, metier_lvl, GearGroup, AreaCode, AreaCode_system, n_trips_sampled, n_trips_total) %>%
#   distinct()
# 
# names(d2O) <- c("Country", "Year", "SamplingScheme", "Gear", "GearGroup", "AreaCode", "AreaCode_system", "NoTripsSampled", "NoTripsTotal")
# d2N <- select(d2div, Country, Year, SamplingScheme, Gear, GearGroup, AreaCode, AreaCode_system, NoTripsSampled, NoTripsTotal) %>%
#   distinct()
# 
# d2final <- rbind(d2O, d2N)


# save(d2div, file = "/ToRc/d2processed.Rdata")
```




```{r}
med <- filter(d2div, AreaCode_system == "GSA")
# Slovenia is grouped with ICES area codes need to fix 
nor <- filter(d2div, !AreaCode_system == "GSA")
```



```{r}
nor1 <- nor %>%
  filter(AreaCode_system == "ICES") %>%
  mutate(
    is_subarea     = str_detect(AreaCode, "^27\\.\\d+$"),
    is_division    = str_detect(AreaCode, "^27\\.\\d+\\.[a-z]$"),
    is_subdivision = str_detect(AreaCode, "^27\\.\\d+\\.[a-z]\\.\\d+")
  ) %>%
  left_join(ices_divisions, by = c("AreaCode" = "subarea")) %>%
  mutate(
    Area_div = case_when(
      is_subarea ~ coalesce(division, AreaCode),  # replace subarea with all its divisions
      TRUE       ~ AreaCode                       # keep divisions/subdivisions as-is
    )
  ) %>%
  select(-division) %>%
  distinct(.orig_row, Area_div, .keep_all = TRUE)


```

```{r, fig.dim= c(12, 8)}

current_year <- year(Sys.Date())

yearsCell <- nor %>%
  filter(Year >= 2017 & Year < 2025) %>% 
  filter(NoTripsSampled != 0 & !Country %in% "EE") %>%
  group_by(AreaCode, GearGroup) %>%
  summarise(NoYears = length(unique(Year)))

df <- yearsCell %>%
  mutate(AreaCode = factor(AreaCode, levels = unique(ices_divisions$division)),
         GearGroup1 = factor(GearGroup, levels = c("DemTrawl", "Gillnetters", "LinesHooks", "PelSeine", "PelTrawl", "Others")))
  
  # aa <- df$AreaCode[is.na(factor(df$AreaCode, levels = unique(ices_divisions$division)))]
  
  
  

ggplot(df, aes(x = GearGroup, y = AreaCode, fill = factor(NoYears))) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "Years") +
  labs(
    x = "Gear",
    y = "ICES area",
    title = "Unique year coverage by gear and ICES area"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r, fig.dim= c(12, 8)}
current_year <- year(Sys.Date())

yearsCell <- med %>%
  filter(Year >= 2017 & Year < 2025) %>% 
  group_by(AreaCode, GearGroup) %>%
  summarise(NoYears = length(unique(Year)))
  
  
ggplot(yearsCell, aes(x = GearGroup, y = AreaCode, fill = factor(NoYears))) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "Years") + 
  labs(
    x = "Gear",
    y = "GSA",
    title = "Unique year coverage by gear and GSA"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}


d4samType <- d4 %>%
  group_by(SamplingScheme,MonitoringMethod) %>%
  summarise(Samname = paste0(unique(SamplingSchemeType), collapse = ","))
# 
# monmeth <- distinct(d4samType, SamplingScheme, MonitoringMethod, SamplingSchemeType)
# 
# 
# dmerged <- left_join(d2div, monmeth, by = c("SamplingScheme"))
# 
# save(dmerged, file = "C:/Users/molla004/OneDrive - Wageningen University & Research/WGCATCH/WGCATCH2025/ToRc/d2merged.Rdata")



```

