---
title: "WGCATCH 2025, ToR c BYC inventory"
author: "Karolina, Sarah"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r}
# PACKAGES
library(tidyverse, quietly = TRUE)
library(DT)
library(readxl)
library(gridExtra)
```

```{r}
# READ data 

# New datacall 
path <- "../ToRc/Data/"
d2 <- read.csv(paste0(path, "D2.csv"))
# Need to filter to avoid overlap 
d2 <- filter(d2, Year > 2020)
d4 <- read.csv(paste0(path, "D4.csv"))

# Old datacall 
d2old <- read_xlsx(paste0(path, "eu.2022.04b_dataset.xlsx"), sheet = "Coverage")
d2old <- filter(d2old, Year >= 2017 & Year <= 2020)
## Remove first row - 
d2old <- d2old[-1, ]
## Rename colnames 
names(d2old) <- c(
  "Country",
  "Year",
  "SamplingScheme",
  "Gear",
  "AreaCode",
  "NoTripsSampled",
  "NoTripsTotal",
  "SamplingIntensity",
  "ImportanceBirds",
  "ImportanceMammals",
  "ImportancePetFish",
  "ImportanceElasmobranchs",
  "ImportanceReptiles",
  "Comments"
)

d4old <- read_xlsx(paste0(path, "eu.2022.04b_dataset.xlsx"), sheet = "Sample schemes bycatch studies")
d4old <- d4old[-1, ]

# RDBES - CE 
ce <- read.csv(paste0(path, "CommercialEffort.csv"))
# Filter out countries that did not respond to the datacall
# unique(d2$Country)
# unique(ce$CEvesselFlagCountry)

ce <- filter(ce, CEvesselFlagCountry %in% unique(d2$Country))
# setdiff(unique(d2$Country), unique(ce$CEvesselFlagCountry)) # med countries and PT-20 excl - correct
```


```{r}
# Useful fun 
classify_gear <- function(x) {
  
  # capture the NAs
  if (is.na(x)) return(NA_character_)
  
  # split on underscores and treat each piece as a separate code
  parts <- unlist(str_split(x, "_"))
  parts <- parts[str_detect(parts, "^[A-Za-z]+$")]
  
  # helper for standalone-ish match
  tok <- function(x) paste0("(?<![A-Z])(", x, ")(?![A-Z])")

  cats <- c(
    PelTrawl = any(str_detect(parts, regex(tok("OTM|PTM|PT"), ignore_case = TRUE))),
    DemTrawl = any(str_detect(parts, regex(tok("OTB|PTB|TBB|OTT|TB"), ignore_case = TRUE))),
    DemSeine = any(str_detect(parts, regex(tok("SSC|SPR|SDN"), ignore_case = TRUE))),
    PelSeine = any(str_detect(parts, regex(tok("PS"), ignore_case = TRUE))),
    LinesHooks = any(str_detect(parts, regex(tok("LLD|LLS|LH|LHP|LHM|LTL"), ignore_case = TRUE))),
    Gillnetters = any(str_detect(parts, regex(tok("GNS|GTR|GN|GND|GTN|GNC"), ignore_case = TRUE)))
  )

  # free-text rules look at full string
  cats["PelTrawl"]    <- cats["PelTrawl"]    | str_detect(x, regex("pelagic trawler", ignore_case = TRUE))
  cats["Gillnetters"] <- cats["Gillnetters"] | str_detect(x, regex("gillnetter", ignore_case = TRUE))
  cats["DemTrawl"]    <- cats["DemTrawl"]    | (str_detect(x, regex("trawler", ignore_case = TRUE)) &
                                                 str_detect(x, regex("demersal|cod", ignore_case = TRUE)))

  present <- names(cats)[cats]

  if (length(present) == 0) {
    "Others"
  } else if (length(present) == 1) {
    present
  } else {
    paste0("Multiple (", paste(sort(present), collapse = "+"), ")")
  }
}

# Expand area
expandIcesArea <- function(x) {
  x <- str_replace_all(x, " ", "")
  x <- str_replace(x, "\\.$", "")   # drop dot

  # Pure 27 region or fully specified already
  if (x == "27") return(x)
  if (str_detect(x, "^27\\.[0-9]+(\\.[A-Za-z0-9]+)*$")) return(x)

  # 27.4b, 27.7bcjk 
  if (str_detect(x, "^27\\.[0-9]+[A-Za-z]+$")) {
    num     <- str_match(x, "^27\\.([0-9]+)")[, 2]
    letters <- str_match(x, "^27\\.[0-9]+([A-Za-z]+)$")[, 2]
    letters_vec <- strsplit(letters, "")[[1]]
    letters_vec <- letters_vec[letters_vec != ""]
    return(sprintf("27.%s.%s", num, letters_vec))
  }

  # 7bcjk, 7fgh etc 
  if (str_detect(x, "^[0-9]+[A-Za-z]+$")) {
    num     <- str_match(x, "^([0-9]+)")[, 2]
    letters <- str_match(x, "^[0-9]+([A-Za-z]+)$")[, 2]
    letters_vec <- strsplit(letters, "")[[1]]
    letters_vec <- letters_vec[letters_vec != ""]
    return(sprintf("27.%s.%s", num, letters_vec))
  }

  # 27.1, 27.12 etc keep as is will be expanded
  if (str_detect(x, "^27\\.\\d+$")) return(x)

  # Fallback: keep unchanged
  x
}

# Parse
parse_one_code <- function(code) {
  orig <- code

  code <- str_squish(code)

  if (is.na(code) || code == "" || code == "NA") {
    return(tibble(
      original       = orig,
      AreaCode_system = NA_character_,
      AreaCode        = NA_character_
    ))
  }

  # Split on both "~" and "," (mixed strings)
  parts <- str_split(code, "[~,]")[[1]] %>%
    str_squish()
  
  parts <- parts[parts != ""]

  map_dfr(parts, function(p) {
    if (is.na(p) || p == "") {
      return(tibble(
        original        = orig,
        AreaCode_system = NA_character_,
        AreaCode        = NA_character_
      ))
    }

   #  explicit GSA + number 
    if (str_detect(str_to_upper(p), "^GSA\\s*\\d+$")) {
      num <- as.integer(str_extract(p, "\\d+"))
      return(tibble(
        original        = orig,
        AreaCode_system = "GSA",
        AreaCode        = paste0("GSA ", num)
      ))
    }

    # Plain digits
    if (str_detect(p, "^\\d{1,2}$")) {
      # If original string has a 27.* and no 'GSA'  interpret as 27.<n> (like 27.12,14)
      if (str_detect(orig, "27\\.") &&
          !str_detect(orig, regex("\\bGSA\\b", ignore_case = TRUE))) {
        new_code <- paste0("27.", p)
        return(tibble(
          original        = orig,
          AreaCode_system = "ICES",
          AreaCode        = new_code
        ))
      } else {
        # Otherwise treat as GSA number
        num <- as.integer(p)
        return(tibble(
          original        = orig,
          AreaCode_system = "GSA",
          AreaCode        = paste0("GSA ", num)
        ))
      }
    }

    p_clean <- str_replace_all(p, " ", "")
    p_clean <- str_replace(p_clean, "\\.$", "")

    if (str_detect(p_clean, "^27(\\.|$)") ||
        str_detect(p_clean, "^[0-9]+[A-Za-z]+$")) {
      expanded <- expandIcesArea(p_clean)
      # expanded may be length > 1
      return(tibble(
        original        = orig,
        AreaCode_system = "ICES",
        AreaCode        = expanded
      ))
    }

    # Unknown stuff 
    tibble(
      original        = orig,
      AreaCode_system = "Unknown",
      AreaCode        = p
    )
  })
}

# 

```

```{r}
# Bind the d2 
d2 <- mutate(d2, Source = "2025DC")
d2old <- mutate(d2old, 
                NoTripsSampled = as.numeric(NoTripsSampled),
                NoTripsTotal = as.numeric(NoTripsTotal),
                Source = "2020DC")
d2all<- bind_rows(d2, d2old)

# mutate PT-20 to PT and the two codes for Germany 
d2all <- mutate(d2all, Country = ifelse(Country %in% "PT-20", "PT", 
                                        ifelse(Country %in% "DEU", "DE", Country)))

```



# Overview

Number of countries responded in the 2020 and 2025 datacall respectively. 

```{r}
d2Tab <- d2all %>%
  group_by(Source) %>%
  summarise(NoCountries = length(unique(Country)),
            Countries = paste0(sort(unique(Country)), collapse = ", "))

d2Tab  %>%
  datatable(extensions = 'Buttons', rownames = FALSE, 
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


# Data cleaning 

## Gear/Metier groupings 

The metiers reported by each country were grouped into six categories: 

1. Pelagic trawl
2. Demersal trawl
3. Demersal seine
4. Pelagic seine
5. Lines and hooks
6. Others 

The table below shows how metiers were grouped to the categories above: 


```{r}
# Group metiers - inventories

d2allg <- d2all %>%
  separate_rows(Gear, sep = "[,;/]\\s*") %>%
  mutate(Gear = str_trim(Gear),
         GearGroup = vapply(Gear, classify_gear, character(1)))
tabgear <- distinct(d2allg, Gear, GearGroup)
tabgear %>%
  datatable(extensions = 'Buttons', rownames = FALSE, 
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
# Group gears - CE - run at the end its taking too long
# ce <- mutate(ce, GearGroup = vapply(CEmetier6, classify_gear, character(1))) # takes too long
# tabgearce <- distinct(ce, CEmetier6, GearGroup)
# datatable(tabgearce, caption = "Grouped metiers in the old datacall")
```


```{r}
# Look up table 
ices_divisions <- tribble(
  ~subarea,   ~division,
  # 27.1
  "27.1",     "27.1.a",
  "27.1",     "27.1.b",
  # 27.2
  "27.2",     "27.2.a",
  "27.2",     "27.2.b",
  # 27.3 (Baltic + Kattegat)
  "27.3.a",   "27.3.a.20",
  "27.3.a",   "27.3.a.21",
  "27.3.b",   "27.3.b.23",
  "27.3.c",   "27.3.c.22",
  # 27.3.d (Baltic Proper) incl. 28.1/28.2
  "27.3.d",   "27.3.d.24",
  "27.3.d",   "27.3.d.25",
  "27.3.d",   "27.3.d.26",
  "27.3.d",   "27.3.d.27",
  "27.3.d",   "27.3.d.28",
  "27.3.d",   "27.3.d.28.1",
  "27.3.d",   "27.3.d.28.2",
  "27.3.d",   "27.3.d.29",
  "27.3.d",   "27.3.d.30",
  "27.3.d",   "27.3.d.31",
  "27.3.d",   "27.3.d.32",
  # 27.4 (North Sea)
  "27.4",     "27.4.a",
  "27.4",     "27.4.b",
  "27.4",     "27.4.c",
  # 27.5 (Iceland & Faroes)
  "27.5",     "27.5.a",
  "27.5",     "27.5.b",
  # 27.6 (W of Scotland/Rockall)
  "27.6",     "27.6.a",
  "27.6",     "27.6.b",
  "27.6.b",   "27.6.b.1",
  # 27.7 (Celtic/Biscay)
  "27.7",     "27.7.a",
  "27.7",     "27.7.b",
  "27.7",     "27.7.c",
  "27.7",     "27.7.d",
  "27.7",     "27.7.e",
  "27.7",     "27.7.f",
  "27.7",     "27.7.g",
  "27.7",     "27.7.h",
  "27.7",     "27.7.j",
  "27.7",     "27.7.k",
  # 27.8 (Bay of Biscay)
  "27.8",     "27.8.a",
  "27.8",     "27.8.b",
  "27.8",     "27.8.c",
  "27.8",     "27.8.d",
  "27.8",     "27.8.e",
  # 27.9
  "27.9",     "27.9.a",
  "27.9",     "27.9.b.1",
  "27.9",     "27.9.b.2",
  "27.9",     "27.9.b",
  # 27.10 (Azores)
  "27.10",    "27.10.a",
  "27.10.a",  "27.10.a.1",
  "27.10.a",  "27.10.a.2",
  "27.10",    "27.10.b",
  # 27.12 (Svalbard/Barents fringe)
  "27.12",    "27.12.a",
  "27.12",    "27.12.b",
  "27.12",    "27.12.c",
  # 27.14 (Greenland)
  "27.14",    "27.14.a",
  "27.14",    "27.14.b"
)
```





```{r}

# Collapse list cols, keep original row
d2e <- d2allg %>%
  ungroup() %>%
  mutate(.orig_row = row_number()) %>%
  rowwise() %>%
  mutate(
    AreaCode_collapsed = paste(AreaCode, collapse = "~")
  ) %>%
  ungroup()

# Expand general
d2_expanded <- d2e %>%
select(-AreaCode) %>%
  mutate(parsed = map(AreaCode_collapsed, parse_one_code)) %>% # the error is in the nested functions above, not here
  unnest(parsed) %>%
  # clean up
  select(-AreaCode_collapsed)

# Expand subarea -> subdiv
d2div <- d2_expanded %>%
  mutate(
    AreaCode = str_trim(AreaCode)   # just in case there are stray spaces
  ) %>%
  left_join(ices_divisions, by = c("AreaCode" = "subarea")) %>%
  mutate(
    # if AreaCode matches a subarea in the lookup, use that division,
    # otherwise keep the original AreaCode
    AreaCode1 = if_else(!is.na(division), division, AreaCode)
  ) %>%
  select(-division) %>%
  distinct(.orig_row, AreaCode1, AreaCode_system, .keep_all = TRUE)

```



```{r}
# Fix specific stuff 
d2div <- mutate(d2div, AreaCode = ifelse(AreaCode %in% "37.2.1", "GSA 17", AreaCode))
# Exclude 3271 France (talked with Sebastien)
# Exclude 27.3.bc 
d2div <- filter(d2div, !(AreaCode %in% "3721" & Country %in% "FR"& Source %in% "2025DC"))
d2div <- filter(d2div, !(AreaCode %in% "27.3.bc" & Country %in% "FR"& Source %in% "2025DC"))
# Reassign 3.a to 4.b (rectangle ovelap issue)
d2div <- mutate(d2div, AreaCode = ifelse(Country %in% "FR" & AreaCode %in% "27.3.a"& Source %in% "2025DC", "27.4.b", AreaCode))

# Exclude Finland area 27.3.d. (all Baltic) - Email for new submission: Done
# d2div <- filter(d2div, !AreaCode %in% "27.3.d.")

# Fix Slovenia 
d2div <- mutate(d2div, AreaCode_system = ifelse(Country %in% "SI", "GSA", AreaCode_system))


# Several records in the old datacall don't have area
naArea <- filter(d2div, is.na(AreaCode))
d2div <- filter(d2div, !is.na(AreaCode))
```





```{r}
d2div <- d2div %>%
  mutate(AreaCode = ifelse(substr(AreaCode1, 1, 4) == "27.3", substr(AreaCode1, 1, 9), sub("^(([^.]*\\.){2}[^.]*).*", "\\1", AreaCode1)))

```


```{r}
ce <- mutate(ce, AreaCode = ifelse(substr(CEarea, 1, 4) == "27.3", substr(CEarea, 1, 9), sub("^(([^.]*\\.){2}[^.]*).*", "\\1", CEarea)))

# TODO - This assumption 
# Filter out records with very low number of trips 
cef <- filter(ce, CEnumberOfFractionalTrips > 0.1)
# quantile(ce$CEnumberOfFractionalTrips, probs = seq(0, 1, by = 0.1), na.rm = TRUE)
# quantile(ce$CEnumberOfDominantTrips, probs = 0.95, na.rm = TRUE)
# hist(ce$CEnumberOfDominantTrips, breaks = 1000, xlim = c(0,200))
# hist(log10(ce$CEofficialDaysAtSea + 0.01), breaks = 100)
# # 
# ggplot(ce, aes(CEnumberOfDominantTrips,CEnumberOfFractionalTrips )) +
#   geom_point()
cef <- distinct(cef, AreaCode, CEmetier6)
cef <- mutate(cef, GearGroup = vapply(CEmetier6, classify_gear, character(1)))

```


```{r}
med <- filter(d2div, AreaCode_system == "GSA")
# Slovenia is grouped with ICES area codes need to fix: Done
nor <- filter(d2div, !AreaCode_system == "GSA")
```

## ICES

```{r}
# Determine sampling schemes with logbooks monitoring method 
d4oldmm <- filter(d4old, `Monitoring method` %in% "Logbook") # old dataset only one and removed either way due to NAs in almost all other variables
d4mm <- filter(d4, MonitoringMethod %in% "LB")  # Estonia
remLB <- unique(d4mm$SamplingScheme)
```


### Coverage by year


```{r, fig.dim= c(16, 12)}

current_year <- year(Sys.Date())

# Factors to be plotted in order 
dfnor <- nor  %>%
  mutate(AreaCode2 = factor(AreaCode, levels = unique(ices_divisions$division)),
         GearGroup1 = factor(GearGroup, levels = c("DemTrawl", "DemSeine","Gillnetters", "LinesHooks", "PelSeine", "PelTrawl", "Others")))

# Remove NAFO areas
dfnor <- filter(dfnor, !is.na(AreaCode2))
  
# With zeros - NAs are created by expanding the df
# These have been left in - Need to check if they correspond to 0 trips sampled
yearsCellWZ <- dfnor %>%
  filter(Year >= 2017 & Year < 2025) %>% # 2017 - 2024
  filter(!SamplingScheme %in% remLB) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoYears = length(unique(Year)))

# Without 0's 
yearsCellWTZ <- dfnor %>%
  filter(Year >= 2017 & Year < 2025) %>% 
  filter(!SamplingScheme %in% remLB) %>%
  filter(NoTripsSampled != 0) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoYears = length(unique(Year)))


a <- ggplot(yearsCellWZ, aes(x = GearGroup1, y = AreaCode2, fill = factor(NoYears))) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "Years") +
  labs(
    x = "Gear",
    y = "ICES area",
    title = "Unique year coverage by gear and ICES area incl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

b <- ggplot(yearsCellWTZ, aes(x = GearGroup1, y = AreaCode2, fill = factor(NoYears))) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "Years") +
  labs(
    x = "Gear",
    y = "ICES area",
    title = "Unique year coverage by gear and ICES area excl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(a, b, ncol=2)

```


### Coverage by sampling scheme 


```{r, fig.dim= c(16, 12)}
# With zeros - NAs are created by expanding the df
# These have been left in - Need to check if they correspond to 0 trips sampled
yearsCellWZ <- dfnor %>%
  filter(Year >= 2017 & Year < 2025) %>% # 2017 - 2024
  filter(!SamplingScheme %in% remLB) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoSampScheme = length(unique(SamplingScheme)))

# Without 0's 
yearsCellWTZ <- dfnor %>%
  filter(Year >= 2017 & Year < 2025) %>% 
  filter(!SamplingScheme %in% remLB) %>%
  filter(NoTripsSampled != 0) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoSampScheme = length(unique(SamplingScheme)))


a <- ggplot(yearsCellWZ, aes(x = GearGroup1, y = AreaCode2, fill = factor(NoSampScheme ))) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "Number of Sampling schemes") +
  labs(
    x = "Gear",
    y = "ICES area",
    title = "Unique sampling schme coverage by gear and ICES area incl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

b <- ggplot(yearsCellWTZ, aes(x = GearGroup1, y = AreaCode2, fill = factor(NoSampScheme ))) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "Number of Sampling schemes") +
  labs(
    x = "Gear",
    y = "ICES area",
    title = "Unique sampling scheme coverage by gear and ICES area excl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(a, b, ncol=2)
```

### Coverage by sampling effort 
```{r, fig.dim= c(16, 12)}
# With zeros - NAs are created by expanding the df
# These have been left in - Need to check if they correspond to 0 trips sampled
yearsCellWZ <- dfnor %>%
  filter(Year >= 2017 & Year < 2025) %>% # 
  filter(!SamplingScheme %in% remLB) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoTripsSampled = sum(NoTripsSampled, na.rm = TRUE))

# Without 0's 
yearsCellWTZ <- dfnor %>%
  filter(Year >= 2017 & Year < 2025) %>% # 
  filter(!SamplingScheme %in% remLB) %>%
  filter(NoTripsSampled != 0) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoTripsSampled = sum(NoTripsSampled, na.rm = TRUE))


a <- ggplot(yearsCellWZ, aes(x = GearGroup1, y = AreaCode2, fill = NoTripsSampled )) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(name = "Number of trips sampled") +
  labs(
    x = "Gear",
    y = "ICES area",
    title = "Unique sampling schme coverage by gear and ICES area incl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

b <- ggplot(yearsCellWTZ, aes(x = GearGroup1, y = AreaCode2, fill = NoTripsSampled)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(name = "Number of trips sampled") +
  labs(
    x = "Gear",
    y = "ICES area",
    title = "Unique sampling scheme coverage by gear and ICES area excl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(a, b, ncol=2)
```

### Comparison with CE

```{r, fig.dim= c(16, 12)}
dfnornew <- filter(dfnor, Source %in% "2025DC") %>%
  filter(Year >= 2021 & Year < 2025) %>% # 
  filter(!SamplingScheme %in% remLB) %>%
  filter(NoTripsSampled != 0) 
cemerg <- left_join(cef, dfnornew, by = c("AreaCode" = "AreaCode2", "GearGroup" = "GearGroup1"))

cefin <- distinct(cemerg, AreaCode, GearGroup, .orig_row) %>%
  mutate(Sampled = ifelse(is.na(.orig_row), "No", "Yes")) %>%
  distinct(AreaCode, GearGroup, Sampled) %>%
   mutate(GearGroup = factor(GearGroup, levels = c("DemTrawl", "DemSeine","Gillnetters", "LinesHooks", "PelSeine", "PelTrawl", "Others")),
          AreaCode = factor(AreaCode, levels = unique(ices_divisions$division)))


ggplot(cefin , aes(x = GearGroup, y = AreaCode, fill = Sampled)) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "SAmpled") +
  labs(
    x = "Gear",
    y = "ICES area",
    title = "Year * gear combination sampled"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```



## Med

### Coverage by year

```{r, fig.dim= c(16, 12)}

# Factors to be plotted in order 
dfmed<- med  %>%
  mutate(AreaCode2 = factor(AreaCode, levels = paste0("GSA ", seq(1:30))),
         GearGroup1 = factor(GearGroup, levels = c("DemTrawl", "DemSeine","Gillnetters", "LinesHooks", "PelSeine", "PelTrawl", "Others")))

# Remove NAFO areas
dfmed <- filter(dfmed, !is.na(AreaCode2))
  
# With zeros - NAs are created by expanding the df
# These have been left in - Need to check if they correspond to 0 trips sampled
yearsCellWZ <- dfmed %>%
  filter(Year >= 2017 & Year < 2025) %>% # 2017 - 2024
  filter(!SamplingScheme %in% remLB) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoYears = length(unique(Year)))

# Without 0's 
yearsCellWTZ <- dfmed %>%
  filter(Year >= 2017 & Year < 2025) %>% 
  filter(!SamplingScheme %in% remLB) %>%
  filter(NoTripsSampled != 0) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoYears = length(unique(Year)))


a <- ggplot(yearsCellWZ, aes(x = GearGroup1, y = AreaCode2, fill = factor(NoYears))) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "Years") +
  labs(
    x = "Gear",
    y = "GSA",
    title = "Unique year coverage by gear and ICES area incl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

b <- ggplot(yearsCellWTZ, aes(x = GearGroup1, y = AreaCode2, fill = factor(NoYears))) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "Years") +
  labs(
    x = "Gear",
    y = "GSA",
    title = "Unique year coverage by gear and ICES area excl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(a, b, ncol=2)



```

### Coverage by sampling scheme 

```{r, fig.dim= c(16, 12)}
# With zeros - NAs are created by expanding the df
# These have been left in - Need to check if they correspond to 0 trips sampled
yearsCellWZ <- dfmed %>%
  filter(Year >= 2017 & Year < 2025) %>% # 2017 - 2024
  filter(!SamplingScheme %in% remLB) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoSampScheme = length(unique(SamplingScheme)))

# Without 0's 
yearsCellWTZ <- dfmed %>%
  filter(Year >= 2017 & Year < 2025) %>% 
  filter(!SamplingScheme %in% remLB) %>%
  filter(NoTripsSampled != 0) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoSampScheme = length(unique(SamplingScheme)))


a <- ggplot(yearsCellWZ, aes(x = GearGroup1, y = AreaCode2, fill = factor(NoSampScheme ))) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "Number of Sampling schemes") +
  labs(
    x = "Gear",
    y = "GSA",
    title = "Unique sampling schme coverage by gear and ICES area incl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

b <- ggplot(yearsCellWTZ, aes(x = GearGroup1, y = AreaCode2, fill = factor(NoSampScheme ))) +
  geom_tile(color = "white") +
  scale_fill_viridis_d(name = "Number of Sampling schemes") +
  labs(
    x = "Gear",
    y = "GSA",
    title = "Unique sampling scheme coverage by gear and ICES area excl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(a, b, ncol=2)
```


### Coverage by sampling effort 

```{r, fig.dim= c(16, 12)}
# With zeros - NAs are created by expanding the df
# These have been left in - Need to check if they correspond to 0 trips sampled
yearsCellWZ <- dfmed %>%
  filter(Year >= 2017 & Year < 2025) %>% # 
  filter(!SamplingScheme %in% remLB) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoTripsSampled = sum(NoTripsSampled, na.rm = TRUE))

# Without 0's 
yearsCellWTZ <- dfmed %>%
  filter(Year >= 2017 & Year < 2025) %>% # 
  filter(!SamplingScheme %in% remLB) %>%
  filter(NoTripsSampled != 0) %>%
  group_by(AreaCode2, GearGroup1) %>%
  summarise(NoTripsSampled = sum(NoTripsSampled, na.rm = TRUE))


a <- ggplot(yearsCellWZ, aes(x = GearGroup1, y = AreaCode2, fill = NoTripsSampled )) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(name = "Number of trips sampled") +
  labs(
    x = "Gear",
    y = "ICES area",
    title = "Unique sampling schme coverage by gear and ICES area incl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

b <- ggplot(yearsCellWTZ, aes(x = GearGroup1, y = AreaCode2, fill = NoTripsSampled)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(name = "Number of trips sampled") +
  labs(
    x = "Gear",
    y = "ICES area",
    title = "Unique sampling scheme coverage by gear and ICES area excl 0 sampled trips"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(a, b, ncol=2)
```





```{r}
d4samType <- d4 %>%
  group_by(SamplingScheme,MonitoringMethod) %>%
  summarise(Samname = paste0(unique(SamplingSchemeType), collapse = ","))

```

