---
title: "Foreign landings overview using the ICES regional database"
author: "Laurent Dubroca, Ruben , Lucia"
date: '`r date()`'
output: pdf_document
bibliography: '/home/moi/datahome/work/biblio/enfin/biblioloran.bib'
---

```{r compil,eval=F,include=F,cache=F,echo=F,warning=F,message=F,progress=F,verbose=F}
#just a reminder for compilation of a docx doc
if(F){
rmarkdown::render("overview.Rmd",c("word_document"))
}
```

*This text should go inside the foreign landings appendix section in the WGCATCH 2018 report*

#Rationale

Foreign landings can be important for some country. They can preclude the
representativness of the sampling plan, because of the difficulties for the 
country to access to the vessels and landings outside its national territory.

To adress this issue, exploratory data analyses are provided based on the fishframe
data format. Two levels of analyses are explored. One at the regional level, where a
quantification of the foreign landings is provided numerically and graphically.
And another at the national level, where a set of table and graphical outputs
are provided to highlight the importance of foreign landings for a
stock, and if needed, to identify the foreign countries where the sampling
effort should be directed.

During the session dedicated to foreign landings in WGCATCH 2018, 8 case study were presented. 
A review of the graphical outputs and the tables with the information used by the participants to diagnose the importance
of the foreign landings in a fishery was done. 
The synthesis of the graphical outputs used is presented in the following table.
The main type of outputs used were barplots and tables, based on landings
aggregated by countries, and the corresponding proportions.

```{r tabplot,eval=T,results="asis"}
library(dplyr)
library(pander)
tabplot<-read.csv("tabplot.csv")
tabplot<-tabplot%>%mutate_all(funs(gsub(", ","\n",.)))
pander(tabplot)
```

#Tools

Analyses and results were generated using the R landuage and environment for
statistical computing [@r2016]. This Rmarkdown report includes text, code and output
in the same document to insure the reproducibility of the analyses and their understanding by end-users [@baumer2014].
The reproducibility of this document is also dependent on the computing environment where the original report
was produced, and can be greatly influenced by the versioning of the software (namely the type of operating system, 
the markdown engine and  R and its packages version).
For example, the extensive use of the `dplyr` package in this report, and the fact that this package is still in
development can lead to different outputs or errors. We let the user to check the compatibility of its own computational environment
and to adapt the code if neede. Information regarding the environment is given at the end of this report.

Working directory, packages loading and general options are set in the following chunk.

```{r global_options,eval=T}
#working directory have to be the directory where is the Rmarkdown file.
#setwd("~/where/is/my/working/directory/") #linux style

#R general option:
options(stringsAsFactors = FALSE)
#knitr chunk option
knitr::opts_chunk$set(cache=FALSE,echo=TRUE, warning=FALSE, 
	message=FALSE, fig.height=6,progress=FALSE,verbose=FALSE,
		include=TRUE,dev='png',autodep=FALSE)
#package
library(dplyr) 
library(ggplot2)
library(maps)
library(mapdata)
library(pander);library(captioner)
library(ggrepel)
library(sf)
library(circlize)
library(tidyr)
library(kableExtra)
#library(DATRAS)
#library(earthmovdist) deprecated
#library(emdist)

#source local functions file
#source("fct.R")
#initialize tables and figures caption
tabcap <- captioner(prefix = "Table")
figcap <- captioner(prefix = "Figure")
```

#Data

The data are formatted in the fishframe format [@ices2016]. Using a standard data format
gives the ability to the countries and the regional
coordination groups (RCG) to
repeat easily the analyses done in this report. Only landings were provided by
ICES to WGCATCH 2018 and consequently the exploratory analyses are based only 
on the commercial landings table (object `CL`).
The information used is the landings agregated by country, metier and year. So
using another source of information for the analyses should
require very little adaptation
A strong hypothese is made regarding the quality of the fishframe data,
rarely used outside the RCG context: in this framework they are considered
representative, without any verification. 

Data are imported in the R workspace.

```{r readdata,eval=T}
#if data.table is not installed, data are read -slowly- with the read.csv base function
if(require(data.table)){
  cl<-data.table::fread("./data/2018 11 01 CL data for WGCATCH 17 species.csv",
		      sep=",",header=TRUE,stringsAsFactors=FALSE,dec=".")
}else{
  cl<-read.csv("./data/2018 11 01 CL data for WGCATCH 17 species.csv")
}
```

##Exploratory data analyses

The data provided cover `r n_distinct(cl$Year)` years of
landings of `r n_distinct(cl$Species)` species for `r n_distinct(cl$FlagCountry)` countries on `r n_distinct(cl$Area)` ICES division.
A summary of the main characteristics of the dataset can be quickly done thanks to the `skim` function.
But in the scope of this document, the corresponding chunk is not evaluated. 

```{r edadata,eval=F}
skimr::skim(cl)%>%kable()#pander()
```

#Regional analyses of foreign landings

Foreign landings can be defined as flows of landing quantity across country.
The corresponding flow matrix is computed for the year 2017, 
by summing the landings by flag country and landed country.
The data are filtered on the year 2017 before processing.

```{r readdata2,eval=T}
#cl<-read.csv("./data/2018 11 01 CL data for WGCATCH 17 species.csv")

pipo<-cl%>%filter(Year==2017)
pipo<-pipo%>%group_by(lanctr=LandingCountry,flagctr=FlagCountry)%>%
		summarise(w=sum(OfficialLandingCatchWeight,na.rm=T)/1000)%>%
		group_by(lanctr)%>%mutate(prop=round(100*w/sum(w)))%>%ungroup()
flowmat<-ggplot(pipo,aes(x=lanctr,y=flagctr,fill=w,label=prop))+
		geom_raster()+
		geom_text()+
		theme_bw()+
                scale_fill_distiller(palette='Spectral',name="Landings (t)",trans="log10")+
        theme(axis.text.x = element_text(size=8, angle=90),
                axis.text.y = element_text(size=8, angle=0))+
		xlab("Landing country")+ylab("Flag country")+ggtitle("Year 2017")
pipo<-pipo%>%spread(lanctr,w,fill=0)
flagctr<-pipo$flagctr
lanctr<-names(pipo)[-1]
pipo<-as.matrix(pipo[,-1])
rownames(pipo) = flagctr
chordDiagram(pipo)
```

#National analyses of foreign landings

The analyses here focus on the foreign landings for a given species for a
country perspective. The case study focus on seabass landings for Netherlands.

Data are first prepared for the analyses.
```{r rub1data,eval=T}
sum <- cl %>%
  filter(Species == "Dicentrarchus labrax", ### fill species here
                     Year == "2017") %>% ### maybe no year filter?
  mutate(origin = if_else(LandingCountry == FlagCountry, "domestic", "foreign"))

dum <- filter(sum, FlagCountry == "NLD") ### fill your country here

tos <- group_by(sum, Year, Area) %>%
  summarise(OfficialLandingCatchWeight = sum(OfficialLandingCatchWeight)) %>%
  mutate(id = "international")

tod <- group_by(dum, Year, Area) %>%
  summarise(OfficialLandingCatchWeight = sum(OfficialLandingCatchWeight)) %>% 
  mutate(id = unique(dum$FlagCountry))
```

## Introduction

`r length(unique(dum$SpeciesDesc))` species (`r unique(dum$SpeciesDesc)`, see http://fishbase.org/Summary/SpeciesSummary.php?ID=63&AT=sea+bass) landings from `r length(unique(dum$FlagCountry))` country in `r length(unique(dum$LandingCountry))` countries are analysed.

## Exploration

```{r rub1explo}
#,echo=FALSE,message=FALSE,warning=FALSE,collapse=TRUE}
full_join(tod, tos) %>% 
  ggplot(aes(Area, OfficialLandingCatchWeight, fill=id)) +
  geom_col(colour = "#000000", position = "dodge") +
  theme_minimal() +
  facet_wrap(~Year)
```



```{r,echo=FALSE,message=FALSE,warning=FALSE,collapse=TRUE}
group_by(dum, Year, Area, LandingCountry) %>%
  summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>%
  ggplot(aes(Year, OfficialLandingCatchWeight)) +
  geom_col(aes(fill=Area),colour="#000000") +
  facet_wrap(~LandingCountry) +
  theme_minimal() +
  labs(caption = paste0(unique(dum$FlagCountry)," commercial landings per Area per Year"))
```


The proportion landings landed in foreign ports are mainly from areas `r as.character((group_by(dum, Area, FlagCountry) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>% arrange(desc(OfficialLandingCatchWeight)) %>% filter(FlagCountry!="domestic"))[1,1])` and `r as.character((group_by(dum, Area, FlagCountry) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>% arrange(desc(OfficialLandingCatchWeight)) %>% filter(FlagCountry!="domestic"))[2,1])`. Whereas OfficialLandingCatchWeightes landed in `r unique(dum$FlagCountry)` are mainly from `r as.character((group_by(dum, Area, FlagCountry) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>% arrange(desc(OfficialLandingCatchWeight)) %>% filter(FlagCountry!="foreign"))[1,1])` and `r as.character((group_by(dum, Area, FlagCountry) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>% arrange(desc(OfficialLandingCatchWeight)) %>% filter(FlagCountry!="foreign"))[2,1])`.



```{r,echo=FALSE,message=FALSE,warning=FALSE,collapse=TRUE}
group_by(dum,FishingActivityCategoryEuropeanLvl6, LandingCountry, Area) %>%
  summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>%
  ggplot(aes(FishingActivityCategoryEuropeanLvl6,OfficialLandingCatchWeight)) +
  geom_col(aes(fill=Area),colour="#000000") +
  labs(caption=paste0(unique(dum$FlagCountry) ," commercial landings per metier per Year")) +
  theme_minimal() +
  facet_wrap(~LandingCountry) +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
```



Main metiers catching `r unique(dum$SpeciesDesc)` are `r as.character((group_by(dum,FishingActivityCategoryEuropeanLvl6) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight))%>% arrange(desc(OfficialLandingCatchWeight)) %>% mutate(prop = round(OfficialLandingCatchWeight/sum(OfficialLandingCatchWeight)*100,2)))[1,1])`, `r as.character((group_by(dum,FishingActivityCategoryEuropeanLvl6) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight))%>% arrange(desc(OfficialLandingCatchWeight)) %>% mutate(prop = round(OfficialLandingCatchWeight/sum(OfficialLandingCatchWeight)*100,2)))[2,1])`, and, `r as.character((group_by(dum,FishingActivityCategoryEuropeanLvl6) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight))%>% arrange(desc(OfficialLandingCatchWeight)) %>% mutate(prop = round(OfficialLandingCatchWeight/sum(OfficialLandingCatchWeight)*100,2)))[3,1])`.



The proportion landings landed in foreign ports are mainly from metiers `r as.character((group_by(dum, FishingActivityCategoryEuropeanLvl6, FlagCountry) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>% arrange(desc(OfficialLandingCatchWeight)) %>% filter(FlagCountry!="Domestic"))[1,1])` and `r as.character((group_by(dum, FishingActivityCategoryEuropeanLvl6, FlagCountry) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>% arrange(desc(OfficialLandingCatchWeight)) %>% filter(FlagCountry!="Domestic"))[2,1])`. Whereas catches landed in `r unique(dum$FlagCountry)` are mainly from `r as.character((group_by(dum, FishingActivityCategoryEuropeanLvl6, FlagCountry) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>% arrange(desc(OfficialLandingCatchWeight)) %>% filter(FlagCountry!="Foreign"))[1,1])` and `r as.character((group_by(dum, FishingActivityCategoryEuropeanLvl6, FlagCountry) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>% arrange(desc(OfficialLandingCatchWeight)) %>% filter(FlagCountry!="Foreign"))[2,1])`.





```{r,echo=FALSE,message=FALSE,warning=FALSE,collapse=TRUE}
group_by(dum,Quarter, LandingCountry, Area) %>%
  summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>%
  ggplot(aes(Quarter,OfficialLandingCatchWeight)) +
  geom_col(aes(fill=Area),colour="#000000") +
  labs(caption=paste0(unique(dum$FlagCountry) ," commercial landings per metier per Year")) +
  theme_minimal() +
  facet_wrap(~LandingCountry) +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
```



```{r,echo=FALSE,message=FALSE,warning=FALSE,collapse=TRUE,eval=F}
kable(group_by(dum,Year, FishingActivityCategoryEuropeanLvl6) %>% summarise(OfficialLandingCatchWeight=sum(OfficialLandingCatchWeight)) %>% mutate(prop = round(OfficialLandingCatchWeight/sum(OfficialLandingCatchWeight)*100,2)) %>% select(-OfficialLandingCatchWeight) %>% spread(Year,prop)) 
```



There are (`r length(unique(dum$FishingActivityCategoryEuropeanLvl6))`) different metiers that land `r unique(dum$SpeciesDesc)`.


#System information

This document was compiled the `r date()` in the following environment.

```{r rversion}
pander(sessionInfo())
```



#Bibliography

